MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-08 chapters/MetaTracts.tex
changed images sizes
changes

%REPLACED BY 
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and future components. The material of choice of a strongly growing number of industrial branches are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands.

CFRP materials show characteristics as high specific stiffness, high specific strength, high corrosion resistance at low weight in comparison to conventional materials. At the same time using CFRPs highly complex and integrated components may be produced, which were previously impossible to manufacture.
Increasingly, primary structures and highly loaded components are being manufactured from CFRP.
Thus CFRPs have successfully been introduced in aeronautic and automotive applications. Besides aerospace and automotive industry, CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.

Carbon fiber reinforced polymer components, more specifically CFRP laminates with endless carbon fibers which we focus on in this work, typically consist of two main components: 
The matrix on the one side, which acts as a bonding component and the reinforcements on the other side, which allows for achieving the desired characteristics e.g. in terms of stiffness, strength or durability. 
To manufacture CFRP laminates with endless carbon fibers, various production processes are in industrial use. 
Most of these processes start with reinforcement component, weaving individual carbon fiber bundles into sheets of carbon fiber cloth in a predefined pattern. 
These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also refereed to as fabric.
Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. 
Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength and stiffness properties. 
Next, resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. 
The main function of this matrix component is to act as a bonding component between the individual carbon fiber bundles. 
For the matrix of CFRP material systems with endless carbon fibers typically resins such as epoxy are frequently used. 
In special applications additional filler materials, such as colors or other particles, may be integrated. 
After curing the resin the production process of the CFRP laminate is finished.

@@

Aside ultrasonic testing, industrial 3D X-Ray computed tomography (XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. 
XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, taken throughout a full rotation of the specimen. 
In industrial XCT, the specimen is placed on the rotary table between X-ray source and detector and penetrated by incident X-rays of the source.
The X-rays pass through the specimen and are attenuated by the material in the specimen.
The detector records 2D projection image depending on its principle, e.g. by transferring the X-rays in a scintillator layer into visible light.  
The specimen is rotated by a predefined angular step and the process is repeated~\cite{heinzl-2008-thesis}. 
In state of the art devices, XCT can reach voxel sizes below $500 nm$.
XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. 
Unfortunately, there is still a trade-off between viewport and image resolution. 
The magnification reached within a cone beam XCT scan is determined by the specified distances
between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport showing more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

In this work we focus on datasets with larger viewports and lower resolutions where the single fibers are not or barely visible. Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of the fiber bundles themselves instead of high resolution studies of individual fibers.
Figure~\ref{fig:data-char} provides an example of our dataset. It also shows an example of the recurring fiber bundle pattern, called a \textit{unit cell}.
Our work is thus motivated from the recent progress in two interrelated fields: On the one hand, fiber reinforced polymer (FRP) components has gained wide application in a variety of industries  because of FRP's superior material and physical properties in comparison to conventional materials \cite{Karpat2012}. On the other hand the use of industrial 3D X-Ray computed tomography (XCT) based non destructive testing of fiber reinforced components has also gained popularity~\cite{Schilling2005}.

@@

While fiber bundles are now understood as increasingly important in determining component properties,
the tools for visualization of the internal structure have not developed at the same pace.
To the best of our knowledge, there is no current work that can resolve simple queries such as:
\begin{itemize}
\item{ How to visualize the geometric structure of a  particular bundle?}
\item{ Which is the largest bundle?}
\item{ Which bundles are in a particular orientation?}
\item{ How to visualize the interaction between a particular pair (crossovers/braiding) or a unit cell?}
\end{itemize}

Extracting the answer to the above queries from the volume rendering of datasets, or from looking at  particular (2D) slices of the data is non-trivial even for experts. We detail and evaluate a method which uses visualization to quickly gain insight into our data. We demarcate the above queries into two, ``geometric structure" and ``spatial context". Geometric structure refers to the shape, size, orientation of a bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other.

We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-Ray computed tomography data of the woven carbon fiber reinforced composites.
The main goal of this work is to expand the state of the art in non destructive testing through visualization of composite structures in ``lower resolutions" and thus ``larger areas" of woven fabrics.

@@

\section {User Evaluation}
\label {sec:evaluation}
\begin{figure}[h]
\centering
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/crop-6-F.PNG}}
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/orientB.PNG}}
\caption{Contextual information: (a) Shows the four largest clusters in the data set in Fig~\ref{fig:prepreg}.
(b)Shows how the red bundle `bends' under the green bundle, which is not immediately apparent from just the volume rendering.}
\label{fig:context}
\end{figure}
\section{Parameter choices}
\label{sec:param_choices}
\begin{figure}
\centering
	\includegraphics[width=0.5\textwidth]{imagesMT2014/crop-16/histogram_10_A.PNG}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top). The unit for length is the grid cube length. Figure~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6.}
	\label{fig:len_dist_crop16}
\end{figure}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions. Which is generally known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$ (Section~\ref{sec:results}). Large $h$ keeps the major bundles intact. For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4 and 5 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements). This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.

>>>>

%REPLACED BY 
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and future components. The material of choice of a strongly growing number of applications are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands in industry. CFRPs have been successfully introduced in aeronautic and automotive applications within recent years, replacing more and more parts made of conventional materials such as aluminum or steel. But besides the aerospace and automotive industry, CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.

In general, CFRP materials show highly desired characteristics as high specific stiffness, high specific strength, high corrosion resistance at considerably lower weight compared to conventional materials. At the same time using CFRPs highly complex and integrated components may be produced, which were previously impossible to manufacture. For example even primary structures and highly loaded components have been increasingly manufactured from CFRP in recent years.

Carbon fiber reinforced polymer components, more specifically CFRP laminates with endless carbon fibers which we focus on in this work, typically consist of two main components: 
The matrix on the one side, which acts as a bonding component and the reinforcements on the other side, which allows for achieving the desired characteristics. To manufacture CFRP laminates with endless carbon fibers, various production processes are in industrial use. Most of these processes start with reinforcement component, weaving individual carbon fiber bundles into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also refereed to as fabric. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength, durability and stiffness properties. Next, resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of this matrix component is to act as a bonding component between the individual carbon fiber bundles. For the matrix of CFRP material systems with endless carbon fibers typically resins such as epoxy are frequently used. In special applications additional filler materials, such as colors or other particles, may be integrated. The production process CFRP laminate is finished after curing the resin and the CFRP laminate is ready for further processing.

@@

Aside ultrasonic testing, industrial 3D X-Ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which taken throughout a full rotation of the specimen. In industrial XCT, the typical scanning geometry foresees the specimen to be placed on the rotary table between X-ray source and detector. X-rays are emitted by the source in the form of a cone beam. When passing through the specimen the X-rays get attenuated by the material in the specimen. The detector records the corresponding 2D attenuation image (penetration image), depending on its principle e.g. by transferring the X-rays in a scintillator layer into visible light. To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}. Finally the series of penetration images is reconstructed to a 3D volumetric dataset of the specimen. 
In state of the art devices, XCT can reach voxel sizes below $500 nm$. XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry. The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but showing more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

In this work we focus on datasets with larger viewports and lower resolutions where the single fibers are not or barely visible. Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of fiber bundles in endless carbon fiber reinforced composites instead of high resolution studies of the individual fibers.
%short fiber reinforced composites as presented by Weissenbock2014 et al.~\cite{Weissenbock2014}.
Figure~\ref{fig:data-char} provides an example of our targeted datasets. It also shows an example of the recurring fiber bundle pattern, called a \textit{unit cell}.
Our work is motivated from the recent progress in two interrelated fields: On the one hand, CFRP components have gained wide application in a variety of industries  because of its superior material and physical properties in comparison to conventional materials \cite{Karpat2012}. On the other hand the use of industrial 3D X-Ray computed tomography (XCT) based non destructive testing of fiber reinforced components has increasingly gained popularity~\cite{Schilling2005}.

@@

While fiber bundles are now understood as highly important in determining component properties, the tools for visualization of the internal structure have not developed at the same pace.
To the best of our knowledge, there is no current work that can resolve simple queries such as:
\begin{itemize}
\item{ How to extract and visualize the geometric structure of a particular bundle?}
\item{ How to visualize the interaction between a particular pair (crossovers/braiding) or a unit cell?}
\item{ Which bundles are in a particular orientation?}
\item{ Which is the largest or smallest bundle?}
\end{itemize}

Extracting the answer to the above queries from the volume rendering of datasets, or from looking at  particular (2D) slices of the data is non-trivial even for experts. We detail and evaluate a method which uses visualization to quickly gain insight into our data. We demarcate the above queries into two parts, ``geometric structure" and ``spatial context". Geometric structure refers to the shape, size, orientation of a bundle or a group of bundles. Spatial context refers to how two or more bundles interact with each other. 
We advance and interpret techniques from diffusion tensor imaging to extract and visualize geometric structures from 3D X-Ray computed tomography data of the woven carbon fiber reinforced composites. The main goal of this work is to expand the state of the art in non destructive testing through visualization of composite structures in ``lower resolutions" and thus ``larger areas" of woven fabrics.

@@

\section {User Evaluation}
\label {sec:evaluation}
\begin{figure}[h]
\centering
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/crop-6-F.PNG}}
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/orientB.PNG}}
\caption{Contextual information: (a) Shows the four largest clusters in the data set in Fig~\ref{fig:prepreg}.
(b)Shows how the red bundle `bends' under the green bundle, which is not immediately apparent from just the volume rendering.}
\label{fig:context}
\end{figure}
\section{Parameter choices}
\label{sec:param_choices}
\begin{figure}
\centering
	\includegraphics[width=0.4\textwidth]{imagesMT2014/crop-16/histogram_10_A.PNG}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top). The unit for length is the grid cube length. Figure~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6. Clusters labeled 2,3,4,5 and 7 all have cardinality less than 10. }
	\label{fig:len_dist_crop16}
\end{figure}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions. Which is generally known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$ (Section~\ref{sec:results}). Large $h$ keeps the major bundles intact. For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4 and 5 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements). This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
